---
title: "Ten Small TidyPanel Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ten Small TidyPanel Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(TidyPanel)
```

The ten bundled CSV files are fictional examples with three input rows each.
They contain no personal records or development datasets.

## Numbers

```{r numbers}
numbers <- read_messy_panel(system.file(
  "extdata", "demo_01_numbers.csv", package = "TidyPanel", mustWork = TRUE
))
numbers
```

The record codes keep their leading zeros. The amount column becomes numeric,
including a negative amount and a missing value.

## Dates

```{r dates}
dates <- read_messy_panel(system.file(
  "extdata", "demo_02_dates.csv", package = "TidyPanel", mustWork = TRUE
))
dates
```

The date column becomes a `Date` column. A missing reading stays missing.

## Subtotals And Periods

```{r subtotals}
panel <- read_messy_panel(
  system.file(
    "extdata", "demo_03_subtotals.csv", package = "TidyPanel", mustWork = TRUE
  ),
  auto_pivot = TRUE,
  return_audit = TRUE
)
panel$data
panel$audit
```

The total row is excluded and the two yearly columns become a long table.

## Percentages

```{r percentages}
percentages <- read_messy_panel(
  system.file("extdata", "demo_04_percentages.csv", package = "TidyPanel", mustWork = TRUE)
)
percentages
```

The rate values become proportions: 0.12, 0.075, and 0.


## Currency

```{r currency}
currency <- read_messy_panel(
  system.file("extdata", "demo_05_currency.csv", package = "TidyPanel", mustWork = TRUE)
)
currency
```

The amount values become 48.25, 16, and -7.5.


## Leading-Zero Codes

```{r codes}
codes <- read_messy_panel(
  system.file("extdata", "demo_06_codes.csv", package = "TidyPanel", mustWork = TRUE)
)
codes
```

Product and shelf codes retain their leading zeros. Quantities become numeric.


## Whitespace

```{r spaces}
spaces <- read_messy_panel(
  system.file("extdata", "demo_07_whitespace.csv", package = "TidyPanel", mustWork = TRUE)
)
spaces
```

Names and labels are trimmed. Numeric values retain their signs and zero.


## Custom Missing Values

```{r missing}
missing <- read_messy_panel(
  system.file("extdata", "demo_08_missing.csv", package = "TidyPanel", mustWork = TRUE),
  na_strings = c("", "NA", "-777")
)
missing
```

Only the caller-selected -777 marker becomes missing. Zero remains a reading.


## Repeated Records

```{r repeated}
repeated <- read_messy_panel(
  system.file("extdata", "demo_09_repeated.csv", package = "TidyPanel", mustWork = TRUE)
)
repeated
```

Both Part A records remain present; cleaning does not automatically deduplicate.


## Clocks And Quality Flags

```{r clocks}
clocks <- read_messy_panel(
  system.file("extdata", "demo_10_clocks_flags.csv", package = "TidyPanel", mustWork = TRUE)
)
clocks
```

Clock text and literal T, F, and M flags are retained without guessing their meaning.


For your own table, supply its file path instead of a bundled example.
