---
title: "Introduction to TidyPanel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to TidyPanel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Welcome to TidyPanel

**TidyPanel** is an industrial-grade parser designed to extract clean, standardized data frames from heavily malformed, human-readable Excel reports. If you have ever struggled to parse financial statements, ERP exports, or complex tables with N-dimensional headers, decoy rows, and embedded subtotals, `TidyPanel` is built for you.

## Features

- **N-Dimensional Header Stitching**: Intelligently identifies multiline headers and forward-fills merged cells to create flat, readable column names.
- **Indentation Hierarchy Extraction**: Uses leading spaces to identify parent-child categorical relationships in accounting and financial reports.
- **Smart Data Amputation**: Automatically removes decorative page breaks, mid-table subtotals, and ghost aggregate rows.
- **Auto Pivot**: Automatically detects temporal columns (e.g., Q1, 2021, FY23) and pivots them into a tidy, long format.
- **Audit Logs**: Full algorithmic transparency via the `return_audit` parameter.

## Basic Usage

To parse an Excel file, simply use the `read_messy_panel()` function.

```{r setup}
# Load the package (simulated for vignette)
# library(TidyPanel)
```

```{r basic_example, eval=FALSE}
# Extract the first valid data block from the spreadsheet
df <- read_messy_panel("data_raw/financial_report.xlsx")

# View the clean data
head(df)
```

## Advanced Features: Hierarchies and Audit Logs

One of TidyPanel's unique features is its ability to understand structure. Financial tables often use indentation to group rows. TidyPanel extracts this structure and places it in a new column called `parent_category`.

Additionally, data parsing should never be a "black box". TidyPanel allows you to generate an **Audit Log** that explicitly records every transformation applied to your data.

```{r audit_example, eval=FALSE}
# Run the parser with auditing turned on
result <- read_messy_panel("data_raw/type_hierarchical_data.xlsx", return_audit = TRUE)

# The cleaned dataframe is in result$data
clean_df <- result$data
head(clean_df)

# The audit log is in result$audit
print(result$audit)
```

**Example Audit Log Output:**

| Operation | Count |
| :--- | :--- |
| Decoy Rows Bypassed | 5 |
| Indentation Hierarchy Extracted | 3 |
| Ghost Bottom Rows Dropped | 4 |

This table guarantees 100% transparency. You can save this log alongside your processed data to maintain a strict audit trail of how the messy raw data was mathematically restructured.

## Cleaning Variable Names

By default, TidyPanel standardizes variable names using the underlying `clean_variable_names()` engine. This function converts text to snake_case, removes special currency/percentage symbols, and maps common synonyms (like `gvkey` or `permno`) to standard terms (`id`).

```{r vars, eval=FALSE}
# Example of variables being cleaned automatically
df <- read_messy_panel("data_raw/messy_headers.xlsx", clean_vars = TRUE)
colnames(df)
```
