---
title: "Three Small TidyPanel Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Three Small TidyPanel Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(TidyPanel)
```

The three bundled CSV files are fictional examples with three input rows each.
They contain no personal records or development datasets.

## Numbers

```{r numbers}
numbers <- read_messy_panel(system.file(
  "extdata", "demo_01_numbers.csv", package = "TidyPanel", mustWork = TRUE
))
numbers
```

The record codes keep their leading zeros. The amount column becomes numeric,
including a negative amount and a missing value.

## Dates

```{r dates}
dates <- read_messy_panel(system.file(
  "extdata", "demo_02_dates.csv", package = "TidyPanel", mustWork = TRUE
))
dates
```

The date column becomes a `Date` column. A missing reading stays missing.

## Subtotals And Periods

```{r subtotals}
panel <- read_messy_panel(
  system.file(
    "extdata", "demo_03_subtotals.csv", package = "TidyPanel", mustWork = TRUE
  ),
  auto_pivot = TRUE,
  return_audit = TRUE
)
panel$data
panel$audit
```

The total row is excluded and the two yearly columns become a long table.

For your own table, supply its file path instead of a bundled example.
